### 2016-02-01 Irregular-occ/Irregular-occ.tex
.

\title{Stellar occultation predictions for  11 irregular satellites of giant planets up to 2017
%\fnmsep\thanks{Table 8 are only available in electronic form at the CDS via anonymous ftp to cdsarc.u-strasbg.fr (130.79.128.5) or via http://cdsweb.u-strasbg.fr/cgi-bin/qcat?J/A+A/ and IAU NSDC data base at www.imcce.fr/nsdc.}
%\fnmsep\thanks{The complete version of Table 8 is available through CDS and IAU NSDC data base at www.imcce.fr/nsdc.}
%\fnmsep\thanks{Partially based on observations made at Laborat\'orio Nacional de Astrof\'{\i}sica (LNA), Itajub\'a-MG, Brazil.}
%\fnmsep\thanks{Partially based on observations through the ESO runs 079.A-9202(A), 075.C-0154, 077.C-0283 and 079.C-0345.}
%\fnmsep\thanks{\textbf{Partially based on observations made at Observatoire de Haute Provence (OHP), F-04870 Saint-Michel l'observatoire, France}}
}

@@

\abstract
{Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size, shape, albedo and composition, to trace back their true origin. The best ground-based technique to determine size and shape, and thus constrain the albedo and in a broader sense composition, is the observation of stellar occultations by these objects.}
{We aim to predict stellar occultations for the eight biggest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda, and for the irregular satellites Phoebe of Saturn and Nereid of Neptune, and also for Triton.}
{We identified candidates to stellar occultations by the irregular satellites from the UCAC4 catalogue and from a catalogue of stars in the sky-path of Neptune obtained from observations made with the ESO2p2/WFI (2.2 m Max-Planck ESO telescope with the Wide Field Imager) instrument. These catalogues were crossed with the ephemeris of the satellites to identify stellar occultations. We used a new ephemeris based solely on the observations from \cite{GomesJunior2015} to generate predictions for the short-time future for the satellites of Jupiter and Phoebe. For Nereid and Triton we used ephemeris from the literature.}
{We managed to identify 396 candidates of stellar occultations between the period of January, 2016 and December, 2017. We made observational tests for the prediction of the event of Himalia in March 03, 2015. The stars and objects were observed close to the date predicted and in the same field to minimize errors and the obtained relative satellite-star positions were used to evaluate the predictions. We call attention for an occultation by Triton on a bright star in October 5, 2017. The event can be observed from Europe and the east coast of USA and will be a good opportunity to study the atmosphere of the satellite.}
{The comparisons between the predictions and the observation tests show a good agreement. We discuss how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.}

\keywords{Occultations - Planets and satellites: general - Planets and satellites: individual: Jovian and Saturnian irregular satellites}

@@

Irregular satellites revolve around giant planets at large distances on eccentric, highly inclined and frequently retrograde orbits. Because of these peculiar orbits, it is largely accepted that these objects did not form by accretion around their planet, but were captured in the early solar system \citep{Sheppard2005}. There is no consensus for a single model explaining where the irregular satellites were formed.

\cite{Cuk2004} showed that the progenitor of the Himalia group may have originated in heliocentric orbits similar to the Hilda asteroid group. \cite{Sheppard2005} stated that the irregular satellites may be some of the objects that were formed within the giant planets region.

@@

\cite{Sheppard2005} and \cite{Jewitt2007} also expose the possibility for the irregular satellites having their origin as comets or TNOs. TNOs are highly interesting objects that, due to their large heliocentric distances, may be highly preserved and have physical properties similar to those they had when they were formed \citep{Barucci2008}. This is even more true for the smaller objects, since in principle larger sizes favour physical differentiation processes in the body and vice-versa. However, due to the distance, the smaller TNOs from this region are more difficult to observe. Thus, if irregular satellites - or at least a few of them - do share a common origin with small TNOs, and since these objects are situated at much closer distances, this gives a unique chance of observing and studying representatives of this specific TNO population in much greater detail than could ever be possible by direct observation of this population in the Kuiper Belt.

@@

In order to obtain precise fundamental physical parameters like size and shape, thus constraining the albedo and in a broader sense also the composition for the irregular satellites and therefore to contribute to the study of their origin, we aim at observing stellar occultations, which provide  more accurate results than other ground-based techniques \citep{Sicardy2011, Ortiz2012, Braga-Ribas2014}.

@@

Triton is an uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but studied together by many authors \citep{Sheppard2005, Jewitt2007}.

Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. However, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007, Elliot_2000}.

No observation of a stellar occultation by an irregular satellite was published up to date, excluding Triton. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have discouraged earlier tries. But, in fact, given their distances to us, current ephemeris and star positions are precise enough for the prediction of the exact location and instant where the shadow of the occultation will cross the Earth. For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds). Thus, in this case, the cumulated uncertainty of ephemeris and star position must be around 40 mas to get good chances of observing a stellar occultation, which is quite feasible today.

\begin{table}
\caption{\label{Tab: satellite-diameter} Estimated diameter of the satellites and correspondent apparent diameter}
\begin{centering}
\begin{tabular}{lccc}
\hline  \hline
\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & mas\tablefootmark{a}  & km & Ref. \tabularnewline
\hline
Ananke & 8 & 29 & 1 \tabularnewline
Carme & 13 & 46 & 1 \tabularnewline
Elara & 24 & 86 & 1 \tabularnewline
Himalia & 41 & $(150\times120) \pm 20$\tablefootmark{b} & 2 \tabularnewline
Leda & 5 & 20 & 1 \tabularnewline
Lysithea & 10 & 36 & 1 \tabularnewline
Pasiphae & 17 & 62 & 1 \tabularnewline
Sinope & 10 & 37 & 1 \tabularnewline
\hdashline
Phoebe & 32 & $212 \pm 1.4$\tablefootmark{b} & 3 \tabularnewline
\hdashline
Nereid & 15 & $340 \pm 50$\tablefootmark{c} & 4 \tabularnewline
Triton & 124 & $2707 \pm 2.0$\tablefootmark{c} & 5 \tabularnewline
\hline
\end{tabular}
\tablebib{
(1) \cite{Rettig2001}; (2) \cite{Porco2003}; (3) \cite{Thomas2010}; (4) \cite{Thomas1991}; (5) \cite{Thomas2000}.}
\end{centering}
\tablefoottext{a}{Using a mean distance from Jupiter of 5 AU, from Saturn of 9 AU and from Neptune of 30 AU.}
\tablefoottext{b}{From Cassini observation.}
\tablefoottext{c}{From Voyager-2 observation.}
\par
\end{table}

@@

Since 2009 many successful observations of stellar occultations by TNOs have been reported in the literature \citep{Elliot2010, Sicardy2011, Ortiz2012, Braga-Ribas2013}, the main disadvantages in their prediction being large distances and ephemeris error, facts somewhat compensated for the larger diameters involved. However, in the case of irregular satellites, we have much better ephemeris because the orbits of their host planets are well known, their observational time span is much wider and covers many orbital periods in contrast to TNOs. Moreover, the irregular satellites are much closer to Earth which implies in a much smaller shadow path error in kilometers. These advantages may be somewhat balanced by the smaller sizes estimated for the irregular satellites.Thus, in comparison, the chances for a successful observation of an stellar occultation by an irregular satellite should be considered at least also as good as that of by a TNO.

@@

%\subsection{Observations used}
In order to have efficient ephemerides for predicting the stellar occultations by irregular satellites, we need predictions based on recent observations.  This is why we decided to develop our own
ephemeris based on the observations published in \cite{GomesJunior2015}. First because many of the authors of this paper worked in the production of that set of positions, so that it was very well known. Second, this consistent set of numerous and precise positions covers many orbital periods at many distinct orbital plane sights, so that the orbital inclinations along with all other orbital elements could be satisfactorily derived without the need of further position sets. For these reasons, only this set of recent positions was used for building the new ephemeris for the satellites.

\cite{GomesJunior2015} published 6523 precise positions for 18 irregular satellites from observations made at the Observatório do Pico dos Dias (OPD), Observatoire Haute-Provence (OHP) and European Southern Observatory (ESO) between 1992 and 2014. %They showed that the orbits of the irregular satellites of the giant planets have systematic errors.
Their observed position offsets relative to the JPL ephemeris reached up to 200 mas for some satellites. At the distance of Jupiter, this represents an difference larger than 700 km in the shadow path of an occultation.%These differences could be associated with errors in their orbital elements.

@@

Our numerical model describes the dynamical evolution of the irregular satellites of Jupiter in a jovicentric reference frame. The satellites are submitted to the influence of the Sun and the rest of the solar system, as well as those of the Galilean satellites and the first harmonics of Jupiter's gravity field. The axis of the reference frame are those of the equatorial reference frame J2000.

We use the following notations: \begin{itemize}
\item $i$ and $l$ one of the irregular satellites of Jupiter
\item $J$ Jupiter 
\item $j$ another body of the Solar System
\item $M_j$ the mass of the $j$ body, not an irregular satellite
\item $m_i$ the mass of the irregular satellite $i$
\item $\vec{r_i}$ the position of the $i$ body with respect to the barycentre of Jupiter System
\item $r_{ij}$ the distance between bodies $i$ and $j$  
\item $R_J$ the radius of Jupiter
\item $J_n$ the dynamic polar oblateness of the nth order for Jupiter's gravity field
\item $U_{\bar{l}\hat{J}}$ potential generated by the oblateness of Jupiter on the satellite $l$
\item $\Phi_i$ is the inclination of the $i$ satellite with respect to Jupiter's equator.
\end{itemize}

@@

\begin{table*}
\caption{Initial osculating elements for Jupiter irregular satellites at JD 2451545.0.% For Themisto's case, the small number of observations meant that the statistical uncertainty was even greater than the obtained elements, though the fitting proved satisfying. We decided to provide its elements with significant number comparable to the other satellite for information, but left it without uncertainty.
 }\label{Tab: sat_ell}
\begin{center}
\begin{tabular}{ccccccc}
\hline\hline
Satellite & a (km) & e & I$\degr$ & $\Omega\degr$ & $\omega\degr$ & $v\degr$ \\ 
\hline
Himalia &   11372100 $\pm$ 500    &    0.166 $\pm$ 0.002      &   45.14 $\pm$ 0.15      &   39.77 $\pm$ 0.19      &   351.48 $\pm$ 0.46      &   97.35 $\pm$ 0.48    \\
Elara &   11741170 $\pm$ 690  &      0.222 $\pm$ 0.002      &   28.64 $\pm$ 0.18      &   68.42 $\pm$ 0.43      &   179.82 $\pm$ 0.56      &   339.08 $\pm$ 0.82  \\
Lysithea &   11739900 $\pm$ 1300  &      0.136 $\pm$ 0.004      &    51.12 $\pm$ 0.27     &   5.53 $\pm$ 0.52      &   53.0 $\pm$ 1.5      &   318.9 $\pm$ 2.0   \\
Leda &   11140300  $\pm$ 4300  &     0.173  $\pm$ 0.007     &   16.15  $\pm$ 0.75    &   272.6  $\pm$ 1.7    &   212.2  $\pm$ 3.6          &   218.8  $\pm$ 3.2  \\
Pasiphae &  23425000  $\pm$ 5000    &     0.379  $\pm$ 0.001       &   152.44 $\pm$ 0.10      &   284.59 $\pm$ 0.21      &   135.96 $\pm$ 0.19      &   236.97 $\pm$ 0.16 \\
Sinope &   22968800 $\pm$ 5200   &     0.316 $\pm$ 0.002      &   157.76 $\pm$ 0.12      &   256.62 $\pm$ 0.55      &   298.38 $\pm$ 0.55      &   167.57 $\pm$ 0.19    \\
Carme &   24202924 $\pm$ 4800      &  0.242 $\pm$ 0.001      &   147.13 $\pm$ 0.10      &   154.01 $\pm$ 0.25      &   47.90 $\pm$ 0.29      &   234.41 $\pm$ 0.19  \\
Ananke &  21683800  $\pm$ 7200  &     0.380 $\pm$ 0.002      &   172.29 $\pm$ 0.20      &   56.9 $\pm$ 1.2      &   123.3 $\pm$ 1.2      &   231.24 $\pm$ 0.21  \\
%Themisto  & 7393800    &     0.198       &   25.77  &   220.0     &   216.3    &   262.1     \\
\hline
\end{tabular} 
\end{center}
\textbf{Notes}: a: Semimajor axis; e : Excentricity; I: Inclination relative to the Equatorial Reference Plane J2000; $\Omega$: Longitude of the ascending node; $\omega$: Argument of periapsis; $v$: True Anomaly.
\end{table*}

All the orbits determined for the satellites show satisfying residuals. %Yet, one of the satellites has a peculiar situation. Its small number of observations in our set means that its orbit is definitely loosely constrained. 
The residuals are lower than those obtained with JPL ephemeris, but the accuracy of an ephemeris decreases when we get further from the time of observations. The main risk of divergence over time comes from the possible absence of long-term effects when fitting to a short timespan of observations. If that were the case, our ephemeris would diverge too quickly to be of any use. JPL ephemerides are fitted over all the available observations. As a result, they will diverge less quickly than our own. Though they are no longer precise enough for our use, they remain a precious reference to identify whether our own model presents a quick divergence.

We compared our ephemeris to the JPL for all the Jupiter satellites we fitted, until 2017. For instance, the divergence between 2015 and 2018 is at most 98 mas in $\Delta \alpha \cos \delta$ and 58 mas in $\Delta \delta$ for Himalia and 181 mas in $\Delta \alpha \cos \delta$ and 152 mas in $\Delta \delta$ for Carme.

In Fig. \ref{Fig: JPL-STE} it is presented the offsets of the positions published by \cite{GomesJunior2015} for the satellite Carme in declination relative to our ephemeris, to \cite{Jacobson2012} JUP300 JPL ephemeris and \cite{Emelyanov2008} ephemeris. It is possible to see that the systematic error point out by \cite{GomesJunior2015} is reduced with our ephemeris, as expected.

\begin{figure}
\includegraphics[width=8.8cm]{figures/Carme_ephemeris.eps} 
\caption{Offsets in declination of the positions published by \cite{GomesJunior2015} for Carme. The red "x" relative to the special-tailored ephemeris, the blue "+" relative to the JUP300 JPL ephemeris and the green dot relative to \cite{Emelyanov2008}. As expected, it is possible to see that the systematic error point out by \cite{GomesJunior2015} is reduced with the STE ephemeris.  \label{Fig: JPL-STE}}
\end{figure}

@@

\begin{figure}
\begin{centering}
\subfigure[SAT375 JPL and PH15]{\includegraphics[width=8.8cm]{figures/Phoebe_JPL-ph15.eps} \label{Fig:eph-Triton-eme}}
\subfigure[\cite{Emelyanov2007} and PH15]{\includegraphics[width=8.8cm]{figures/Phoebe_EME-ph15.eps} \label{Fig:eph-Triton-zhang}}
\caption{Comparison between the PH15, SAT375 JPL and \cite{Emelyanov2007} ephemeris for the satellite Phoebe.}
\label{Fig:eph-Phoebe}
\end{centering}
\end{figure}

@@

For Triton, we used the most recent ephemeris published by \cite{Emelyanov2015}. In Fig. \ref{Fig:eph-Triton} we compare the ephemeris with the ephemeris published in \cite{Zhang2014} and the NEP081 JPL ephemeris \citep{Jacobson2009}. It is possible to see that the difference between the three ephemeris is smaller than 15 mas for the period 2015-2018. This value is smaller than the apparent size of Triton (124 mas, see Table \ref{Tab: satellite-diameter}), which means a good probability of observing an event by this object.

@@

For Nereid, we used the ephemeris published by \cite{Emelyanov2011}, which is more recent than the JPL ephemeris published by \cite{Jacobson2009}. The comparison between the two ephemeris can be seen in Fig. \ref{Fig:eph-Nereid}. The difference between them is smaller than 60 mas which is bigger than the satellite apparent diameter (see Table \ref{Tab: satellite-diameter}). Although this fact, the difference is bigger in Right Ascension than in Declination, which means that, in an event, the error would be smaller in the latitude where the shadow will pass on Earth than the error at the time of the occultation.

@@

@@

The prediction of the occultations was made by crossing the stellar coordinates and proper motions of the UCAC4 catalogue \citep{Zacharias2013} with the ephemeris explicited in Sec. \ref{Sec: integration}. The search for stellar candidates follows the same procedure as presented by \cite{Assafin2010, Assafin2012} and \cite{Camargo2014}.

We predicted occultation for the 8 major irregular satellites of Jupiter:  Ananke, Carme, Elara, Himalia, Leda, Lysithea, Pasiphae and Sinope; Phoebe of Saturn and Triton and Nereid of Neptune.

>>>>

\title{Stellar occultation predictions for  10 irregular satellites of giant planets plus Triton for 2016-2017
%\fnmsep\thanks{Table 8 are only available in electronic form at the CDS via anonymous ftp to cdsarc.u-strasbg.fr (130.79.128.5) or via http://cdsweb.u-strasbg.fr/cgi-bin/qcat?J/A+A/ and IAU NSDC data base at www.imcce.fr/nsdc.}
%\fnmsep\thanks{The complete version of Table 8 is available through CDS and IAU NSDC data base at www.imcce.fr/nsdc.}
%\fnmsep\thanks{Partially based on observations made at Laborat\'orio Nacional de Astrof\'{\i}sica (LNA), Itajub\'a-MG, Brazil.}
%\fnmsep\thanks{Partially based on observations through the ESO runs 079.A-9202(A), 075.C-0154, 077.C-0283 and 079.C-0345.}
%\fnmsep\thanks{\textbf{Partially based on observations made at Observatoire de Haute Provence (OHP), F-04870 Saint-Michel l'observatoire, France}}
}

@@

\abstract
{Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size, shape, albedo and composition, to trace back their true origin. The best ground-based technique to determine size and shape, and thus constrain the albedo and in a broader sense composition, is the observation of stellar occultations by these objects.}
{We aim to predict stellar occultations for the eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda, and for the irregular satellites Phoebe of Saturn and Nereid of Neptune, and also for Triton.}
{We identified candidates to stellar occultations by the irregular satellites from the UCAC4 catalogue and from a catalogue of stars in the sky-path of Neptune obtained from observations made with the ESO2p2/WFI (2.2 m Max-Planck ESO telescope with the Wide Field Imager) instrument. These catalogues were crossed with the ephemeris of the satellites to identify stellar occultations. We used a new ephemeris based solely on the observations from \cite{GomesJunior2015} to generate predictions for the short-time future for the satellites of Jupiter and Phoebe. For Nereid and Triton we used recently published ephemeris.}
{We managed to identify 396 candidates of stellar occultations between the period of January, 2016 and December, 2017. We made observational tests for the prediction of the event of Himalia in March 03, 2015. The stars and objects were observed close to the date predicted and in the same field to minimize errors and the obtained relative satellite-star positions were used to evaluate the predictions. We call attention for an occultation by Triton of a bright star (R=12.5) in October 5, 2017. The event can be observed from Europe and the east coast of USA and will be a good opportunity to access the current state of the atmosphere of the satellite.}
{The comparisons between the predictions and the observation tests show a good agreement. We discuss how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.}

\keywords{Occultations - Planets and satellites: general - Planets and satellites: individual: Jovian and Saturnian irregular satellites - Planets and satellites: individual: Triton}

@@

Irregular satellites revolve around giant planets at large distances in eccentric, highly inclined and frequently retrograde orbits. Because of these peculiar orbits, it is largely accepted that these objects did not form by accretion around their planet, but were captured in the early solar system \citep{Sheppard2005}.

There is no consensus for a single model explaining where the irregular satellites were formed. \cite{Cuk2004} showed that the progenitor of the Himalia group may have originated in heliocentric orbits similar to the Hilda asteroid group. \cite{Sheppard2005} stated that the irregular satellites may be some of the objects that were formed within the giant planets region.

@@

\cite{Sheppard2005} and \cite{Jewitt2007} also explored the possibility that the irregular satellites originated as comets or TNOs. TNOs are highly interesting objects that, due to their large heliocentric distances, may be highly preserved with physical properties similar to those they had when they were formed \citep{Barucci2008}. This is even more true for the smaller objects, since in principle larger sizes favour physical differentiation processes in the body and vice-versa. However, due to the distance, the smaller TNOs from this region are more difficult to observe. Thus, if irregular satellites - or at least a few of them - do share a common origin with small TNOs, and since these objects are situated at much closer heliocentric distances now, this gives a unique chance of observing and studying representatives of this specific TNO population in much greater detail than could ever be possible by direct observation of this population in the Kuiper Belt.

@@

In order to obtain precise fundamental physical parameters like size and shape, thus constraining the albedo and in a broader sense also the composition for the irregular satellites and therefore to contribute to the study of their origin, we aim at observing stellar occultations, which provide  more accurate results than other ground-based techniques \citep{Sicardy2011, Ortiz2012, Braga-Ribas2014}. For that, reliable predictions of stellar occultations by these satellites are most needed.

@@

Triton is an uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but still studied together by many authors \citep{Sheppard2005, Jewitt2007}. Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. However, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007, Elliot_2000}. For all these reasons, Triton was also included as a target in this work.

Excluding Triton \citep{Olkin1997, Elliot_2000}, no observation of a stellar occultation by an irregular satellite was published up to date. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have discouraged earlier tries. But, in fact, given their relatively closer distances as compared to TNOs and Centaurs, and considering the current precision of their ephemeris and of star positions, we can now reliably predict the exact location and instant where the shadow of the occultation will cross the Earth. For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds). Thus, in this case, if the accumulated error (2 sigma or 95\% confidence level) of ephemeris and star position be around 70 mas, we have a probability of about 30\% of observing a stellar occultation, which is quite satisfactory today (see discussion in section \ref{Sec: discussion}).

\begin{table}
\caption{\label{Tab: satellite-diameter} Estimated diameter of the satellites and correspondent apparent diameter}
\begin{centering}
\begin{tabular}{lccc}
\hline  \hline
\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & mas\tablefootmark{a}  & km & Ref. \tabularnewline
\hline
Ananke & 8 & 29 & 1 \tabularnewline
Carme & 13 & 46 & 1 \tabularnewline
Elara & 24 & 86 & 1 \tabularnewline
Himalia & 41 & $(150\times120) \pm 20$\tablefootmark{b} & 2 \tabularnewline
Leda & 5 & 20 & 1 \tabularnewline
Lysithea & 10 & 36 & 1 \tabularnewline
Pasiphae & 17 & 62 & 1 \tabularnewline
Sinope & 10 & 37 & 1 \tabularnewline
\hdashline
Phoebe & 32 & $212 \pm 1.4$\tablefootmark{b} & 3 \tabularnewline
\hdashline
Nereid & 15 & $340 \pm 50$\tablefootmark{c} & 4 \tabularnewline
Triton & 124 & $2707 \pm 2.0$\tablefootmark{c} & 5 \tabularnewline
\hline
\end{tabular}
\tablebib{
(1) \cite{Rettig2001}; (2) \cite{Porco2003}; (3) \cite{Thomas2010}; (4) \cite{Thomas1991}; (5) \cite{Thomas2000}.}
\end{centering}
\tablefoottext{a}{Using a mean distance from Jupiter of 5 AU, from Saturn of 9 AU and from Neptune of 30 AU.}
\tablefoottext{b}{From Cassini observations.}
\tablefoottext{c}{From Voyager-2 observations.}
\par
\end{table}

@@

Since 2009 many successful observations of stellar occultations by TNOs have been reported in the literature \citep{Elliot2010, Sicardy2011, Ortiz2012, Braga-Ribas2013}, the main disadvantages in their prediction being large heliocentric distances and ephemeris error, facts somewhat compensated for the larger diameters involved. However, in the case of irregular satellites, we have much better ephemeris because the orbits of their host planets are well known, their observational time span is much wider and covers many orbital periods in contrast to TNOs. Moreover, the irregular satellites are much closer to Earth which implies in a much smaller shadow path error in kilometers. These advantages may be somewhat balanced by the smaller sizes estimated for the irregular satellites.Thus, in comparison, the chances for a successful observation of an stellar occultation by an irregular satellite should be considered at least also as good as those by TNOs.

@@

%\subsection{Observations used}
In order to improve ephemeris in short time spans in the near future for a more efficient prediction of stellar occultations by irregular satellites, we need predictions based on new and recent observations. In this context, \cite{GomesJunior2015} published 6523 precise positions for 18 irregular satellites from observations made at the Observatório do Pico dos Dias (OPD), Observatoire Haute-Provence (OHP) and European Southern Observatory (ESO) between 1992 and 2014. %They showed that the orbits of the irregular satellites of the giant planets have systematic errors.
%Their observed position offsets relative to the JPL ephemeris reached up to 200 mas for some satellites. At the distance of Jupiter, this represents an difference larger than 700 km in the shadow path of an occultation.%These differences could be associated with errors in their orbital elements.

Here we develop our own ephemeris based on the observations published in \cite{GomesJunior2015}. First because many of the authors of this paper worked in the production of that set of positions, so that it was very well known. Second, besides recent observations, this consistent set of numerous and precise positions covers many orbital periods at many distinct orbital plane sights, so that the orbital inclinations along with all other orbital elements could be satisfactorily derived without the need of further position sets. For these reasons, only this set of positions was used for building the new ephemeris for the satellites of Jupiter.

@@

Our numerical model describes the dynamical evolution of the irregular satellites of Jupiter in a jovicentric reference frame. The satellites are submitted to the influence of the Sun and the rest of the solar system, as well as that of the Galilean satellites and the first harmonics of Jupiter's gravity field. The axes of the reference frame are those of the ICRS.

We use the following notations: \begin{itemize}
\item $i$ and $l$ one of the irregular satellites of Jupiter
\item $J$ Jupiter 
\item $j$ another body of the Solar System
\item $M_j$ the mass of the $j$-th body, not an irregular satellite
\item $m_i$ the mass of the irregular satellite $i$
\item $\vec{r_i}$ the position of the $i$-th body with respect to the barycentre of Jupiter System
\item $r_{ij}$ the distance between bodies $i$ and $j$  
\item $R_J$ the radius of Jupiter
\item $J_n$ the dynamic polar oblateness of the n-th order for Jupiter's gravity field
\item $U_{\bar{l}\hat{J}}$ potential generated by the oblateness of Jupiter on the satellite $l$
\item $\Phi_i$ is the inclination of the $i$-th satellite with respect to Jupiter's equator.
\end{itemize}

@@

\begin{table*}
\caption{Initial osculating elements for Jupiter irregular satellites at JD 2451545.0.% For Themisto's case, the small number of observations meant that the statistical uncertainty was even greater than the obtained elements, though the fitting proved satisfying. We decided to provide its elements with significant number comparable to the other satellite for information, but left it without uncertainty.
 }\label{Tab: sat_ell}
\begin{center}
\begin{tabular}{ccccccc}
\hline\hline
Satellite & a (km) & e & I$\degr$ & $\Omega\degr$ & $\omega\degr$ & $v\degr$ \\ 
\hline
Himalia &   11372100 $\pm$ 500    &    0.166 $\pm$ 0.002      &   45.14 $\pm$ 0.15      &   39.77 $\pm$ 0.19      &   351.48 $\pm$ 0.46      &   97.35 $\pm$ 0.48    \\
Elara &   11741170 $\pm$ 690  &      0.222 $\pm$ 0.002      &   28.64 $\pm$ 0.18      &   68.42 $\pm$ 0.43      &   179.82 $\pm$ 0.56      &   339.08 $\pm$ 0.82  \\
Lysithea &   11739900 $\pm$ 1300  &      0.136 $\pm$ 0.004      &    51.12 $\pm$ 0.27     &   5.53 $\pm$ 0.52      &   53.0 $\pm$ 1.5      &   318.9 $\pm$ 2.0   \\
Leda &   11140300  $\pm$ 4300  &     0.173  $\pm$ 0.007     &   16.15  $\pm$ 0.75    &   272.6  $\pm$ 1.7    &   212.2  $\pm$ 3.6          &   218.8  $\pm$ 3.2  \\
Pasiphae &  23425000  $\pm$ 5000    &     0.379  $\pm$ 0.001       &   152.44 $\pm$ 0.10      &   284.59 $\pm$ 0.21      &   135.96 $\pm$ 0.19      &   236.97 $\pm$ 0.16 \\
Sinope &   22968800 $\pm$ 5200   &     0.316 $\pm$ 0.002      &   157.76 $\pm$ 0.12      &   256.62 $\pm$ 0.55      &   298.38 $\pm$ 0.55      &   167.57 $\pm$ 0.19    \\
Carme &   24202924 $\pm$ 4800      &  0.242 $\pm$ 0.001      &   147.13 $\pm$ 0.10      &   154.01 $\pm$ 0.25      &   47.90 $\pm$ 0.29      &   234.41 $\pm$ 0.19  \\
Ananke &  21683800  $\pm$ 7200  &     0.380 $\pm$ 0.002      &   172.29 $\pm$ 0.20      &   56.9 $\pm$ 1.2      &   123.3 $\pm$ 1.2      &   231.24 $\pm$ 0.21  \\
%Themisto  & 7393800    &     0.198       &   25.77  &   220.0     &   216.3    &   262.1     \\
\hline
\end{tabular} 
\end{center}
\textbf{Notes}: a: semimajor axis; e : excentricity; I: inclination relative to the equatorial reference plane J2000; $\Omega$: longitude of the ascending node; $\omega$: argument of periapsis; $v$: true anomaly.
\end{table*}

All the orbits determined for the satellites show satisfying residuals. %Yet, one of the satellites has a peculiar situation. Its small number of observations in our set means that its orbit is definitely loosely constrained. 
The residuals are smaller than those obtained with JPL ephemeris, which was expected because the accuracy of an ephemeris decreases when we get further from the time of observations. The main risk of divergence over time comes from the possible absence of long-term effects when fitting to a short timespan of observations. If that were the case, our ephemeris would diverge too quickly to be of any use. JPL ephemerides are fitted over all the available observations. As a result, they will diverge less quickly than our own. Though they are no longer precise enough for our use, they remain a precious reference to identify whether our own model presents a quick divergence.

We compared our ephemeris to the JPL for all the Jupiter satellites we fitted, until 2018. For instance, the divergence between 2015 and 2018 is at most 98 mas in $\Delta \alpha \cos \delta$ and 58 mas in $\Delta \delta$ for Himalia and 181 mas in $\Delta \alpha \cos \delta$ and 152 mas in $\Delta \delta$ for Carme.

Fig. \ref{Fig: JPL-STE} displays the offsets of the positions published by \cite{GomesJunior2015} for the satellite Carme in declination relative to our ephemeris, to \cite{Jacobson2012} JUP300 JPL ephemeris and \cite{Emelyanov2008} ephemeris.We see that the systematic JPL ephemeris offsets pointed out by \cite{GomesJunior2015} are reduced with our ephemeris, as expected.

\begin{figure}
\includegraphics[width=8.8cm]{figures/Carme_ephemeris.eps} 
\caption{Offsets in declination of the positions published by \cite{GomesJunior2015} for Carme. The red "x" relative to the special-tailored ephemeris, the blue "+" relative to the JUP300 JPL ephemeris and the green dot relative to \cite{Emelyanov2008}. As expected, the ephemeris systematic errors pointed out by \cite{GomesJunior2015} are reduced with the STE ephemeris.  \label{Fig: JPL-STE}}
\end{figure}

@@

\begin{figure}
\begin{centering}
\includegraphics[width=8.8cm]{figures/Phoebe.eps} 
\caption{Comparison between the PH15, SAT375 JPL and \cite{Emelyanov2007} ephemeris for the satellite Phoebe.}
\label{Fig:eph-Phoebe}
\end{centering}
\end{figure}

@@

For Triton, we used the most recent ephemeris published by \cite{Emelyanov2015}. In Fig. \ref{Fig:eph-Triton} we compare it with the ephemeris published in \cite{Zhang2014} and the NEP081 JPL ephemeris \citep{Jacobson2009}. The offsets between the three ephemeris are smaller than 15 mas for the period 2015-2018. This value is much smaller than the apparent size of Triton (124 mas, see Table \ref{Tab: satellite-diameter}), which indicates a good probability of observing an event by this object.

@@

For Nereid, we used the ephemeris published by \cite{Emelyanov2011}, which uses more recent observations than the JPL ephemeris published by \cite{Jacobson2009}. The comparison between the two ephemeris can be seen in Fig. \ref{Fig:eph-Nereid}. The right ascension offsets between them are smaller than 60 mas and the declination ones are smaller than 15 mas. For Nereid, ephemeris errors in right ascension translate to uncertainties in the central instant of a stellar occultation, so offsets of 60 mas are not critical. The declination ephemeris offsets are of the order of the apparent diameter (15 mas, see Table \ref{Tab: satellite-diameter}), resulting in fair chances (50\%) that the shadow crosses the expected Earth latitude predicted for an occultation. % which is bigger than the satellite apparent diameter (see Table \ref{Tab: satellite-diameter}). Although this fact, the difference is bigger in Right Ascension than in Declination, which means that, in an event, the error would be smaller in the latitude where the shadow will pass on Earth than the error at the time of the occultation.

@@

\cite{GomesJunior2015} published 902 new positions for Nereid from observations between 1992 and 2014. The development of a better ephemeris for Nereid using these new positions in the context discussed here of near future stellar occultations is desirable, but is out of the scope of this paper.

@@

The prediction of the occultations was made by crossing the stellar coordinates and proper motions of the UCAC4 catalogue \citep{Zacharias2013} with the ephemeris presented in Sec. \ref{Sec: integration}. The search for stellar candidates follows the same procedure as presented by \cite{Assafin2010, Assafin2012} and \cite{Camargo2014}.

We predicted occultations for the 8 major irregular satellites of Jupiter,  Ananke, Carme, Elara, Himalia, Leda, Lysithea, Pasiphae and Sinope, and for Phoebe Phoebe of Saturn and Triton and Nereid of Neptune.
